Stability in human interaction networks: primitive typology of vertex, prominence of measures and time activity statistics

### 2015-04-15 paper.tex
recauchutada revisao com estatistica circular e +++

Next subsection exposes related work. Section~\ref{sec:data} details data while Section~\ref{sec:carac} is dedicated to the characterization methods. Section~\ref{sec:results} holds results and discussions, leading to Section~\ref{sec:conc} for conclusions and further work. Appendix~\ref{ap:ded} develops simple graph results used in the methods, Appendix~\ref{scripts} exposes access to data and scripts, Appendix~\ref{ap:circular} presents the circular statistics (or directional statistics) used in characterizing time activity, and Appendix~\ref{ap:typ} further introduces the typological approach. Detailed tables and figures are bundled in Support Information, and are cited upon need.

@@

  \subsection{Time activity statistics}                          
  Messages were counted along time with respect to seconds, minutes, hours, days of the week, days of the month, and months of the year. This resulted in histograms from which obvious patterns can be observed. Furthermore, the ration $\frac{b_h}{b_l}$ of the highest and lowest incidences on the histograms served as a hint of how the distribution relates to the uniform distribution. Mean and standard deviations in time were taken using circular statistics, as described in Appendix~\ref{ap:circular}.

@@

In other words, let $\sigma$ and $\phi$ be integers in the intervals $1 \leq \sigma \leq 6$, $1 \leq \phi \leq 3$, and the basic six Erd\"os sectioning possibilities $\{E_{\sigma}\}$ have three Erd\"os sectors $E_{\sigma}= \{e_{\sigma, \phi} \}$ defined as:

\begin{alignat}{3}\label{eq:part}
e_{\sigma,1}&=\{\;i\;|\;\overline{k}_{\sigma,\Leftarrow}\geq&&\overline{k}_{\sigma,i}\} \nonumber \\
e_{\sigma,2}&=\{\;i\;|\;\overline{k}_{\sigma,\Leftarrow}<\;&&\overline{k}_{\sigma,i}\leq\overline{k}_{\sigma,\Rightarrow}\} \\ 
e_{\sigma,3}&=\{\;i\;|\;&&\overline{k}_{\sigma,i}<\overline{k}_{\sigma,\Rightarrow}\} \nonumber
\end{alignat}

\noindent where $\{\overline{k}_{\sigma,i}\}$ is:

@@

\noindent and both $\overline{k}_{\sigma,\Leftarrow}$ and $\overline{k}_{\sigma,\Rightarrow}$ are found using $P(\overline{k})$ or $\hat{P}(\overline{k})$ as described above.

@@

%\begin{alignat}{3}
\begin{equation}
\begin{split}
%\begin{multline}
C_1&=\{c_{1,\phi}=\left\{i\mid i\;\in e_{\sigma,\phi}, \;\forall\; \sigma\}\right\} \\
C_2&=\{c_{2,\phi}=\left\{i\mid \exists \;\;\sigma : i \in e_{\sigma,\phi}\}\right\} \\
C_3&=\{c_{3,\phi}=\left\{i\mid i\;\in e_{\sigma,\phi'}, \;\forall\; \sigma,\;\forall\;\phi'\geq \phi\}\right\} \\
C_4&=\{c_{4,\phi}=\left\{i\mid i\;\in e_{\sigma,\phi'}, \;\forall\; \sigma,\;\forall\;\phi'\leq \phi\}\right\} \\
C_5&=\{c_{5,\phi}=\left\{i\mid i\;\in e_{\sigma,\phi'}, \;\forall\; \sigma,\right.\\
	  &\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \left.\;\forall\;(\phi'+1)\%4\leq (\phi+1)\%4\}\right\} \\
C_6&=\{c_{6,\phi}=\left\{i\mid i\;\in e_{\sigma,\phi'}, \;\forall\; \sigma,\right.\\
	  &\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \left.\;\forall\;(\phi'+1)\%4\geq (\phi+1)\%4\}\right\} \\
%\end{multline}
\end{split}
\end{equation}
%\end{alignat}

@@

\subsection{Constancy and discrepancy of activity along time}\label{constDisc}
One remarkable feature from the analysis is that the activity 
along time is practically the same for all lists. Circular mean and standard deviation
for each time scale is given in Table~\ref{tab:circ}.

\begin{table*}[t]
	\caption{Circular mean, standard deviation, variance, circular dispersion and the relation of maximum and minimum incidence at each time unit. The Appendix~\ref{ap:circular} exposes the theoretical background considered of directional (or circular) statistics.}
\begin{center}
    \begin{tabular}{ |l|| c|c|c|c|c| }
        \hline
	scale & mean & standard deviation & variance & circular dispersion & $\frac{max(incidence)}{min(incidence)}$ \\ \hline\hline
	\input{tables/tab2TimeLAU}
    \end{tabular}
\end{center}
\label{tab:circ}
\end{table*}

@@

\begin{figure*} 
   \centering
        \includegraphics[width=\textwidth,height=12cm]{figs/im13PCAPLOT}
	\caption{Prevalence of symmetry measures over clustering for data dispersion. This was considered an important result of PCA analysis, together with stability on component formation. First plot holds the usual degree versus clustering coefficient layout. Second plot is similar, but the first component is an average of centrality measures (see Table~\ref{compPCA}). Third plot resulted from applying PCA with symmetry measures (see Table~\ref{compPCA2}), where greater dispersion can be observed in all sectors, although more pronounced in intermediary and hubs sectors.
    For this figure, a window size of $ws = 1000$ messages was used from LAU list. The general layout is recurrent and consistent with the literature: most connected vertices have low clusterization while higher clusterization is gradually more incident as the number of connections is lowered.
Similar structures were observed in all window sizes $ws\;\in\;[500,10000]$ and for networks of other email lists, which exposes a common relation held by degree, strength and betweenness measures to clustering coefficient.
    In the third plot, clustering coefficient is omitted as it is only relevant for third component. In this case, second component is representative of symmetry measurements of vertex interactions. Dispersion suggests symmetry related measures are more powerful for characterizing interaction networks than clustering coefficient, specially for hubs and intermediary vertices.
    }
    \label{fig:sym}
\end{figure*}

@@

    \subsection{Primitive typology from Erd\"os sectors}\label{sec:pty}
Consider the sector to which a participant belongs as her/his type. General considerations of this typology are:
\begin{itemize}
	\item A participant belongs to many networks (his family, family of a friend, the email list to which he belongs, work friends, etc.).
	\item A participant might belong to all three sectors at the same time. Actually, it is reasonable to assume that almost every person belongs to all three sectors for some of the networks.
	\item It is given by construction that a participant is of just one type at a network in a given snapshot (exception for inclusive compound criteria given in Section~\ref{sectioning}).
	\item The participant often transitions from one sector to another within a network.
	\item Reported stability of network structure arises from activity with continuous change of the participant types.
	\item Type often changes with the scale of the activity window. For example, two snapshots of 200 and 10000 messages will probably reveal participants with more than one type. 
\end{itemize}

@@

@@

\subsection{Third party libraries and software}
The programming framework used 
is mainly Python-based, with emphasis on usual
scientific tools. More specifically,
scripts where written for 2.7.3 version of Python,
with the following third party libraries: Numpy, Pylab/Matplotlib, NetworkX, IGraph.
Behind the scenes, Graphviz is accessed via PyGraphviz to make network drawings.
\section{Circular statistics}\label{ap:circular}
\section{Typologies}\label{ap:typ}
There are other ways to split and characterize networks. To point a common example, the center of the network is defined as all the nodes whose maximum distance to any other node is the radius (the radius is the minimum maximum distance to all vertices , i.e. the radius is the minimum eccentricity). 
In the same framework, the periphery (as opposed to the center) consists of the nodes whose maximum distance to any node is the diameter (diameter being the maximum geodesic on the network). Accordingly, the intermediary sector can be defined as the nodes that are not in the center or in the periphery. Interestingly, in the email networks analyzed, with such criteria, the center can often be a factor of 4 times larger than the periphery and the intermediary group often exceeds 90\% of the nodes~\cite{networkx}.

\section{Tables}\label{sectables}
\clearpage
\subsection{PCA tables}\label{sec:pcat}

\begin{table}[H]
  \centering
  \caption{Principal components composition in the simplest case: with degree, clustering coefficient and betweenness centrality. LAU list, $ws=1000$ messages in 20 disjoint positioning was used for statistics. The first component is a weighted average of degree and betweenness centrality. The second component is mostly clustering coefficient. The first and second components represent more than 95\% of total variance. The $\lambda$ bottom line holds the percentage of total variance attributed to each component.}
  \begin{tabular}{|l|c|c| c|c| c|c|}\hline
 & \multicolumn{2}{c|}{PC1}          & \multicolumn{2}{c|}{PC2} & \multicolumn{2}{c|}{PC3}  \\\hline
       & $\mu$            & $\sigma$ & $\mu$         & $\sigma$ & $\mu$ & $\sigma$  \\\hline
$d$       & {\bf 48.02}   & 1.39     & 2.82          & 1.74     & 48.09  & 0.32 \\
$cc$      & 4.12          & 2.94     & {\bf 90.45}   & 3.98     & 3.98  & 0.77 \\ 
$bt$      & {\bf 47.87}   & 1.55     & 6.74          & 4.08     & 47.93 & 0.46 \\ \hline
$\lambda$ & 64.67         & 0.52     & 33.26         & 0.23     & 2.08  & 0.40 \\ \hline
  \end{tabular}
  \label{compPCA0}
\end{table}

\begin{table}
  \centering
  \caption{Principal components composition in percentages. LAU list, $ws=1000$ messages in 20 disjoint positioning was used for statistics. First component is a weighted average of degree and strength and betweenness centrality. The second component is mostly related to the  clustering coefficient. The first and second components represent more than 90\% of the variance.}
  \begin{tabular}{|l|c|c| c|c| c|c|}\hline
 & \multicolumn{2}{c|}{PC1} & \multicolumn{2}{c|}{PC2} & \multicolumn{2}{c|}{PC3}  \\\hline
       & $\mu$ & $\sigma$ & $\mu$ & $\sigma$ & $\mu$ & $\sigma$  \\\hline
$d$       & {\bf 14.58} & 0.14 & 0.43  & 0.35 & 1.51  & 1.08 \\
$d^{in}$  & {\bf 14.12} & 0.14 & 1.71  & 1.22 & 17.80 & 6.20 \\
$d^{out}$ & {\bf 13.95} & 0.12 & 2.80  & 1.83 & 21.15 & 5.62 \\
$s$       & {\bf 14.48} & 0.13 & 0.78  & 0.65 & 5.51  & 4.71 \\ 
$s^{in}$  & {\bf 14.10} & 0.14 & 2.17  & 1.28 & 17.32 & 6.11 \\ 
$s^{out}$ & {\bf 14.05} & 0.13 & 2.08  & 1.14 & 19.31 & 4.86 \\ \hline
$cc$      & 0.99        & 0.70 & {\bf 83.38} & 4.83 & 2.75  & 1.62 \\ 
$bt$      & {\bf 13.73} & 0.19 & 6.65  & 1.31 & 14.66 & 10.14 \\ \hline
$\lambda$ & 81.80 & 0.83 & 12.53 & 0.09  & 3.24 & 0.62 \\ \hline
  \end{tabular}
  \label{compPCA}
\end{table}

\begin{table}
  \centering
  \caption{Principal components formation with symmetry-related metrics (see Section~\ref{measures}). LAU list, $ws=1000$ messages in 20 disjoint positioning was used for statistics. In this case, clusterization is pushed to the third principal component. The second component is primarily derived from symmetry measurements, but also out-degree and out-strength, and disequilibrium standard deviation. Betweenness centrality again has a role similar to degree, but weaker. The clusterization component combines with disequilibrium, while asymmetry is combined to out-degree and out-strength. The three components have in average 80.36\% of the variance.}
  \begin{tabular}{|l|c|c| c|c| c|c|}\hline
 & \multicolumn{2}{c|}{PC1} & \multicolumn{2}{c|}{PC2} & \multicolumn{2}{c|}{PC3}  \\\hline
       & $\mu$ & $\sigma$ & $\mu$ & $\sigma$ & $\mu$ & $\sigma$  \\\hline
$d$            & {\bf 11.51} & 0.42 & 2.00  & 0.76 & 2.39  & 0.49 \\ 
$d^{in}$       & {\bf 11.45} & 0.34 & 2.86  & 0.91 & 1.68  & 0.67 \\
$d^{out}$      & {\bf 10.68} & 0.60 & {\bf 7.43}  & 1.00 & 3.00  & 1.02 \\
$s$            & {\bf 11.37} & 0.42 & 1.75  & 0.71 & 4.31  & 0.63 \\ 
$s^{in}$       & {\bf 11.33} & 0.35 & 2.39  & 1.10 & 3.69  & 0.86 \\ 
$s^{out}$      & {\bf 10.74} & 0.55 & {\bf 6.14}  & 1.05 & 4.75  & 0.98 \\ \hline
$cc$           & 0.91        & 0.64 & 2.68  & 1.67 & {\bf 22.27} & 6.43 \\ 
$bt$           & {\bf 10.87} & 0.38 & 1.17  & 0.93 & 4.03  & 1.42 \\ \hline
$asy$          & 3.99        & 1.45 & {\bf 18.13} & 1.67 & 2.55  & 1.77 \\
$\mu_{asy}$    & 4.15        & 1.40 & {\bf 17.07} & 1.78 & 2.49  & 1.67 \\
$\sigma_{asy}$ & 1.21        & 0.67 & {\bf 17.49} & 0.79 & 3.29  & 2.33 \\
$dis$          & 5.78        & 0.51 & 1.94  & 1.28 & {\bf 24.75} & 3.73 \\
$\mu_{dis}$    & 0.79        & 0.49 & {\bf 14.00} & 1.14 & 3.73  & 3.13 \\
$\sigma_{dis}$ & 5.18        & 0.72 & 4.93  & 2.48 & {\bf 17.04} & 4.78 \\ \hline
$\lambda$      & 51.09       & 1.07 & 20.04 & 1.31 & 9.23  & 6.63 \\ \hline
  \end{tabular}
  \label{compPCA2}
\end{table}

\clearpage

@@

\begin{table}
    \caption{Distribution of activity among agents. First column is dedicated to percentage of messages sent by the most active participant. Column for the first quartile ($1Q$) exhibits minimum percentage of participants responsible for at least 25\% of total messages. Similarly, the column for the first three quartiles $1-3Q$ exhibits minimum percentage of participants responsible for 75\% of total messages. The last decile $10D$ column has maximum percentage of participants responsible for 10\% of messages.}
\begin{center}
    \begin{tabular}{ | l ||  c | c | c | c | }
        \hline
        list & hub & $ 1Q $ & $ 1-3Q $ & $10D$ \\ \hline
    CPP & 14.41 & 0.19 (27.8\%) & 4.09 (75.13\%) & 83.65 (-10.04\%)   \\ 
    MET & 11.14 & 0.81 (30.61\%) & 8.33 (75,11\%) & 80.49 (-10.02\%)  \\ 
    LAU & 2.78 & 1.10 (25.16\%) & 13.02 (75,04\%) & 67.37 (-10.03\%) \\ 
    LAD & 4.00 & 0.95 (25.50\%) & 11.83 (75,07\%) & 71.13 (-10.03\%) \\\hline
    \end{tabular}
\end{center}
\label{autores}
\end{table}

>>>>

Next subsection exposes related work. Section~\ref{sec:data} details data while Section~\ref{sec:carac} is dedicated to the characterization methods. Section~\ref{sec:results} holds results and discussions, leading to Section~\ref{sec:conc} for conclusions and further work. Appendix~\ref{ap:ded} develops simple graph results used in the methods, Appendix~\ref{scripts} exposes access to data and scripts, and Appendix~\ref{ap:typ} further introduces the typological approach. Detailed tables and figures are bundled in Support Information, and are cited upon need.

@@

\subsection{Time activity statistics}\label{sec:mtime}
  Messages were counted along time with respect to seconds, minutes, hours, days of the week, days of the month, and months of the year. This resulted in histograms from which obvious patterns can be observed. Furthermore, the ration $\frac{b_h}{b_l}$ of the highest and lowest incidences on the histograms served as a hint of how the distribution relates to the uniform distribution.

  Measures of mean and dispersion were taken using circular statistics. For such, each $measure$ (data point) is represented as a complex number of unit magnitude $z=e^{i\theta}=\cos(\theta)+i\sin(\theta)$, where $\theta=measure\frac{2\pi}{period}$. Moments $m_n$, lengths of moments $R_n$, mean angle $\theta_\mu$, and rescaled mean angle $\theta_\mu'$ are defined as:

\begin{align}\label{eq:cmom}
	m_n&=\frac{1}{N}\sum_{i=1}^N z_i^n \nonumber\\
	R_n&=|m_n|\\
	\theta_\mu&=Arg(m_1) \nonumber \\
	\theta_\mu'&=\frac{period}{2\pi} \theta_\mu \nonumber
\end{align}

The $\theta_\mu'$ is used as the measure of location. Dispersion is measured using the circular variance $Var(z)$, the circular standard deviation $S(z)$, and the circular dispersion $\delta(z)$:

\begin{align}\label{eq:cmd}
	Var(z)&=1 - R_1 \nonumber\\
	S(z)&= \sqrt{-2\ln(R_1)}\\
	\delta(z)&=\frac{1-R_2}{2 R_1^2} \nonumber
\end{align}

Although very simple, such measures enables both comparison of the email networks exposed in this article, and the deliverance of a benchmark to which other interaction networks can be compared.

@@

In other words, let $\gamma$ and $\phi$ be integers in the intervals $1 \leq \gamma \leq 6$, $1 \leq \phi \leq 3$, and the basic six Erd\"os sectioning possibilities $\{E_{\gamma}\}$ have three Erd\"os sectors $E_{\gamma}= \{e_{\gamma, \phi} \}$ defined as:

\begin{alignat}{3}\label{eq:part}
e_{\gamma,1}&=\{\;i\;|\;\overline{k}_{\gamma,L}\geq&&\overline{k}_{\gamma,i}\} \nonumber \\
e_{\gamma,2}&=\{\;i\;|\;\overline{k}_{\gamma,L}<\;&&\overline{k}_{\gamma,i}\leq\overline{k}_{\gamma,R}\} \\ 
e_{\gamma,3}&=\{\;i\;|\;&&\overline{k}_{\gamma,i}<\overline{k}_{\gamma,R}\} \nonumber
\end{alignat}

\noindent where $\{\overline{k}_{\gamma,i}\}$ is:

@@

\noindent and both $\overline{k}_{\gamma,L}$ and $\overline{k}_{\gamma,R}$ are found using $P(\overline{k})$ or $\hat{P}(\overline{k})$ as described above.

@@

%\begin{alignat}{3}
\begin{equation}
\begin{split}
%\begin{multline}
C_1&=\{c_{1,\phi}=\left\{i\mid i\;\in e_{\gamma,\phi}, \;\forall\; \gamma\}\right\} \\
C_2&=\{c_{2,\phi}=\left\{i\mid \exists \;\;\gamma: i \in e_{\gamma,\phi}\}\right\} \\
C_3&=\{c_{3,\phi}=\left\{i\mid i\;\in e_{\gamma,\phi'}, \;\forall\; \gamma,\;\forall\;\phi'\geq \phi\}\right\} \\
C_4&=\{c_{4,\phi}=\left\{i\mid i\;\in e_{\gamma,\phi'}, \;\forall\; \gamma,\;\forall\;\phi'\leq \phi\}\right\} \\
C_5&=\{c_{5,\phi}=\left\{i\mid i\;\in e_{\gamma,\phi'}, \;\forall\; \gamma,\right.\\
	  &\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \left.\;\forall\;(\phi'+1)\%4\leq (\phi+1)\%4\}\right\} \\
C_6&=\{c_{6,\phi}=\left\{i\mid i\;\in e_{\gamma,\phi'}, \;\forall\; \gamma,\right.\\
	  &\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \left.\;\forall\;(\phi'+1)\%4\geq (\phi+1)\%4\}\right\} \\
%\end{multline}
\end{split}
\end{equation}
%\end{alignat}

@@

\subsection{Constancy and discrepancy of activity along time}\label{constDisc}
One remarkable feature from the analysis is that the activity 
along time is practically the same for all lists. Circular mean and standard deviation
for each time scale are given in Table~\ref{tab:circ}.

\begin{table*}[t]
	\caption{The rescaled circular mean $\theta_\mu'$, the standard deviation $S(z)$, the variance $Var(z)$, the circular dispersion $\delta(z)$ and the relation of maximum and minimum incidence at each time unit $\frac{max(incidence}{min(incidence}$. The Section~\ref{sec:mtime} exposes the theoretical background considered of directional (or circular) statistics.}
\begin{center}
    \begin{tabular}{ |l|| c|c|c|c|c| }
        \hline
	scale & $\theta_\mu'$ & $S(z)$ & $Var(z)$ & $\delta(z)$ & $\frac{max(incidence)}{min(incidence)}$ \\ \hline\hline
	\input{tables/tab2TimeLAU}
    \end{tabular}
\end{center}
\label{tab:circ}
\end{table*}

@@

\begin{figure*} 
   \centering
        \includegraphics[width=.6\textwidth,height=10cm]{figs/im13PCAPLOT}
	\caption{Prevalence of symmetry measures over clustering for data dispersion. This was considered an important result of PCA analysis, together with stability on component formation. First plot holds the usual degree versus clustering coefficient layout. Second plot is similar, but the first component is an average of centrality measures (see Table~\ref{compPCA}). Third plot resulted from applying PCA with symmetry measures (see Table~\ref{compPCA2}), where greater dispersion can be observed in all sectors, although more pronounced in intermediary and hubs sectors.
    For this figure, a window size of $ws = 1000$ messages was used from LAU list. The general layout is recurrent and consistent with the literature: most connected vertices have low clusterization while higher clusterization is gradually more incident as the number of connections is lowered.
Similar structures were observed in all window sizes $ws\;\in\;[500,10000]$ and for networks of other email lists, which exposes a common relation held by degree, strength and betweenness measures to clustering coefficient.
    In the third plot, clustering coefficient is omitted as it is only relevant for third component. In this case, second component is representative of symmetry measurements of vertex interactions. Dispersion suggests symmetry related measures are more powerful for characterizing interaction networks than clustering coefficient, specially for hubs and intermediary vertices.
    }
    \label{fig:sym}
\end{figure*}

@@

    \subsection{Primitive typology from Erd\"os sectors}\label{sec:pty}
Consider the sector to which a participant belongs as her/his type. General considerations of this typology are:
\begin{itemize}
	\item A participant belongs to many networks (his family, family of a friend, the email list to which he belongs, work friends, etc.).
	\item A participant might belong to all three sectors at the same time. Actually, it is reasonable to assume that almost every person belongs to all three sectors for some of the networks.
	\item It is given by construction that a participant is of just one type at a network in a given snapshot (exception for the compound inclusive and exclusive criterion given in Section~\ref{sectioning}).
	\item The participant often transitions from one sector to another within a network.
	\item Reported stability of network structure arises from activity with continuous change of the participant types.
	\item Type often changes with the scale of the activity window. For example, two snapshots of 200 and 10000 messages will probably reveal participants with more than one type. 
\end{itemize}

@@

\section{Typologies}\label{ap:typ}
There are other ways to split and characterize networks. To point a common example, the center of the network is defined as all the nodes whose maximum distance to any other node is the radius (the radius is the minimum maximum distance to all vertices , i.e. the radius is the minimum eccentricity). 
In the same framework, the periphery (as opposed to the center) consists of the nodes whose maximum distance to any node is the diameter (diameter being the maximum geodesic on the network). Accordingly, the intermediary sector can be defined as the nodes that are not in the center or in the periphery. Interestingly, in the email networks analyzed, with such criteria, the center can often be a factor of 4 times larger than the periphery and the intermediary group often exceeds 90\% of the nodes~\cite{networkx}.

@@

\subsection{Third party libraries and software}
The programming framework used 
is mainly Python-based, with emphasis on usual
scientific tools. More specifically,
scripts where written for 2.7.3 version of Python,
with the following third party libraries: Numpy, Pylab/Matplotlib, NetworkX, IGraph.
Behind the scenes, Graphviz is accessed via PyGraphviz to make network drawings.
\clearpage
\section{Tables}\label{sectables}
\subsection{PCA tables}\label{sec:pcat}
\begin{table}[!h]
  \centering
  \caption{Principal components composition in the simplest case: with degree, clustering coefficient and betweenness centrality. LAU list, $ws=1000$ messages in 20 disjoint positioning was used for statistics. The first component is a weighted average of degree and betweenness centrality. The second component is mostly clustering coefficient. The first and second components represent more than 95\% of total variance. The $\lambda$ bottom line holds the percentage of total variance attributed to each component.}
  \begin{tabular}{|l|c|c| c|c| c|c|}\hline
 & \multicolumn{2}{c|}{PC1}          & \multicolumn{2}{c|}{PC2} & \multicolumn{2}{c|}{PC3}  \\\hline
       & $\mu$            & $\sigma$ & $\mu$         & $\sigma$ & $\mu$ & $\sigma$  \\\hline
$d$       & {\bf 48.02}   & 1.39     & 2.82          & 1.74     & 48.09  & 0.32 \\
$cc$      & 4.12          & 2.94     & {\bf 90.45}   & 3.98     & 3.98  & 0.77 \\ 
$bt$      & {\bf 47.87}   & 1.55     & 6.74          & 4.08     & 47.93 & 0.46 \\ \hline
$\lambda$ & 64.67         & 0.52     & 33.26         & 0.23     & 2.08  & 0.40 \\ \hline
  \end{tabular}
  \label{compPCA0}
\end{table}

\begin{table}[!h]
  \centering
  \caption{Principal components composition in percentages. LAU list, $ws=1000$ messages in 20 disjoint positioning was used for statistics. First component is a weighted average of degree and strength and betweenness centrality. The second component is mostly related to the  clustering coefficient. The first and second components represent more than 90\% of the variance.}
  \begin{tabular}{|l|c|c| c|c| c|c|}\hline
 & \multicolumn{2}{c|}{PC1} & \multicolumn{2}{c|}{PC2} & \multicolumn{2}{c|}{PC3}  \\\hline
       & $\mu$ & $\sigma$ & $\mu$ & $\sigma$ & $\mu$ & $\sigma$  \\\hline
$d$       & {\bf 14.58} & 0.14 & 0.43  & 0.35 & 1.51  & 1.08 \\
$d^{in}$  & {\bf 14.12} & 0.14 & 1.71  & 1.22 & 17.80 & 6.20 \\
$d^{out}$ & {\bf 13.95} & 0.12 & 2.80  & 1.83 & 21.15 & 5.62 \\
$s$       & {\bf 14.48} & 0.13 & 0.78  & 0.65 & 5.51  & 4.71 \\ 
$s^{in}$  & {\bf 14.10} & 0.14 & 2.17  & 1.28 & 17.32 & 6.11 \\ 
$s^{out}$ & {\bf 14.05} & 0.13 & 2.08  & 1.14 & 19.31 & 4.86 \\ \hline
$cc$      & 0.99        & 0.70 & {\bf 83.38} & 4.83 & 2.75  & 1.62 \\ 
$bt$      & {\bf 13.73} & 0.19 & 6.65  & 1.31 & 14.66 & 10.14 \\ \hline
$\lambda$ & 81.80 & 0.83 & 12.53 & 0.09  & 3.24 & 0.62 \\ \hline
  \end{tabular}
  \label{compPCA}
\end{table}

\begin{table}[!h]
  \centering
  \caption{Principal components formation with symmetry-related metrics (see Section~\ref{measures}). LAU list, $ws=1000$ messages in 20 disjoint positioning was used for statistics. In this case, clusterization is pushed to the third principal component. The second component is primarily derived from symmetry measurements, but also out-degree and out-strength, and disequilibrium standard deviation. Betweenness centrality again has a role similar to degree, but weaker. The clusterization component combines with disequilibrium, while asymmetry is combined to out-degree and out-strength. The three components have in average 80.36\% of the variance.}
  \begin{tabular}{|l|c|c| c|c| c|c|}\hline
 & \multicolumn{2}{c|}{PC1} & \multicolumn{2}{c|}{PC2} & \multicolumn{2}{c|}{PC3}  \\\hline
       & $\mu$ & $\sigma$ & $\mu$ & $\sigma$ & $\mu$ & $\sigma$  \\\hline
$d$            & {\bf 11.51} & 0.42 & 2.00  & 0.76 & 2.39  & 0.49 \\ 
$d^{in}$       & {\bf 11.45} & 0.34 & 2.86  & 0.91 & 1.68  & 0.67 \\
$d^{out}$      & {\bf 10.68} & 0.60 & {\bf 7.43}  & 1.00 & 3.00  & 1.02 \\
$s$            & {\bf 11.37} & 0.42 & 1.75  & 0.71 & 4.31  & 0.63 \\ 
$s^{in}$       & {\bf 11.33} & 0.35 & 2.39  & 1.10 & 3.69  & 0.86 \\ 
$s^{out}$      & {\bf 10.74} & 0.55 & {\bf 6.14}  & 1.05 & 4.75  & 0.98 \\ \hline
$cc$           & 0.91        & 0.64 & 2.68  & 1.67 & {\bf 22.27} & 6.43 \\ 
$bt$           & {\bf 10.87} & 0.38 & 1.17  & 0.93 & 4.03  & 1.42 \\ \hline
$asy$          & 3.99        & 1.45 & {\bf 18.13} & 1.67 & 2.55  & 1.77 \\
$\mu_{asy}$    & 4.15        & 1.40 & {\bf 17.07} & 1.78 & 2.49  & 1.67 \\
$\sigma_{asy}$ & 1.21        & 0.67 & {\bf 17.49} & 0.79 & 3.29  & 2.33 \\
$dis$          & 5.78        & 0.51 & 1.94  & 1.28 & {\bf 24.75} & 3.73 \\
$\mu_{dis}$    & 0.79        & 0.49 & {\bf 14.00} & 1.14 & 3.73  & 3.13 \\
$\sigma_{dis}$ & 5.18        & 0.72 & 4.93  & 2.48 & {\bf 17.04} & 4.78 \\ \hline
$\lambda$      & 51.09       & 1.07 & 20.04 & 1.31 & 9.23  & 6.63 \\ \hline
  \end{tabular}
  \label{compPCA2}
\end{table}

@@

\begin{table}[!h]
    \caption{Distribution of activity among agents. First column is dedicated to percentage of messages sent by the most active participant. Column for the first quartile ($1Q$) exhibits minimum percentage of participants responsible for at least 25\% of total messages. Similarly, the column for the first three quartiles $1-3Q$ exhibits minimum percentage of participants responsible for 75\% of total messages. The last decile $10D$ column has maximum percentage of participants responsible for 10\% of messages.}
\begin{center}
    \begin{tabular}{ | l ||  c | c | c | c | }
        \hline
        list & hub & $ 1Q $ & $ 1-3Q $ & $10D$ \\ \hline
    CPP & 14.41 & 0.19 (27.8\%) & 4.09 (75.13\%) & 83.65 (-10.04\%)   \\ 
    MET & 11.14 & 0.81 (30.61\%) & 8.33 (75,11\%) & 80.49 (-10.02\%)  \\ 
    LAU & 2.78 & 1.10 (25.16\%) & 13.02 (75,04\%) & 67.37 (-10.03\%) \\ 
    LAD & 4.00 & 0.95 (25.50\%) & 11.83 (75,07\%) & 71.13 (-10.03\%) \\\hline
    \end{tabular}
\end{center}
\label{autores}
\end{table}
